---
title: "p8105_hw1_mbc2178"
author: "Melvin Coleman"
date: "2022-09-23"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (tidyverse)
```

## Problem 1 

```{r}
data("penguins", package = "palmerpenguins")
```

```{r}
# The penguins dataset is printed below and consists of the following varies including species, island, bill length etc...

head(penguins, 5)
```
There are `r nrow(penguins)` observations (rows) and `r ncol(penguins)` variables (columns). 

The mean flipper length in the `penguins` dataset is `r round(mean(pull(penguins, flipper_length_mm), na.rm=TRUE))` measured in millimeters (mm). 

## Scatter Plot of Flipper Length vs Bill Length by Species

Next, we made a scatter plot of `flipper_length_mm` vs `bill_length_mm` by penguin
`species` below and saved the scatter plot to my project directory. 

```{r scatterplot}
plot_peng <- ggplot(penguins, aes(x = bill_length_mm,
                                  y = flipper_length_mm, color = species)) + 
             geom_point()

ggsave("penguins.pdf", height = 4, width = 6)
```

## Problem 2 

```{r, create_df}

peng_df <- tibble(
  norm_samp = rnorm (10, mean = 0),
  
  norm_samp_pos = norm_samp > 0,
  
  vec_hw = c("basketball", "handball", "streetball", "soccer","netball", 
              "dodgeball","football","kickball","softball","pongball"), 
  
 vec_factor_hw = factor(c("MARRIED","WIDOWED","SINGLE","MARRIED","SINGLE","WIDOWED","SINGLE","SINGLE","WIDOWED"
                            ,"MARRIED"),levels = c("MARRIED", "WIDOWED", "SINGLE"))
)
```
The mean of variable `norm_samp` is `r mean(pull(peng_df,norm_samp))`.  

The mean of `norm_samp_pos` is `r mean(pull(peng_df, norm_samp_pos))`.

The mean of `vec_hw` is `r mean(pull(peng_df, vec_hw))`.

The mean of `vec_factor_hw` is `r mean(pull(peng_df, vec_factor_hw))` 

I noticed that trying to take the mean of the variables `vec_hw` and 
`vec_factor_hw` in the `peng_df` dataset yielded invalid results. In short, 
R couldn't perform the operation because the argument was not numeric or 
logical.

Nevertheless,I was able to calculate the mean of  variable `norm_samp` and 
`norm_samp_pos` in this dataset. 

Moving on, now let's apply the `as.numeric` function to our logical variable `norm_samp_pos`,
character variable `vec_hw`, and factor variable `vec_factor_hw`. 

```{r convert_var, eval = FALSE}
  mean(as.numeric(pull(peng_df, norm_samp_pos)))
  mean(as.numeric(pull(peng_df, vec_hw)))
  mean(as.numeric(pull(peng_df, vec_factor_hw)))
```

After applying the `as.numeric` function to the logical, character, and factor variables, I noticed that I still received a mean value for my logical variable
and factor variable. This was because R treated the "TRUE" and "FALSE" in my logical variable as 1 and 0 respectively. 

In my factor variable, R also treated the three levels (MARRIED, SINGLE and WIDOWED) as numeric when this function was applied. Therefore, each factor level 
was given numeric codes corresponding to the alphabetical order in the vector. 

For the vector character variable, when R tried to convert it to numeric it couldn't and therefore returned "NA". 

This helps to explain why we couldn't take the mean of the character and factor 
variable in the first code we ran earlier.















