---
title: "p8105_hw1_mbc2178"
author: "Melvin Coleman"
date: "2022-09-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (tidyverse)
```

## Problem 1 

```{r}
data("penguins", package = "palmerpenguins")
```

The section below provides a brief description of the penguins dataset. 
The penguin dataset consists of the following summary statistics:
`r summary(penguins)` .
In this dataset, there are `r nrow(penguins) ` observations and 
`r ncol(penguins`) variables. 
The mean flipper length in the penguins dataset is `r mean(pull(penguins, flipper_length_mm), na.rm=TRUE)` measured in millimeters(mm). 

## Scatter Plot of Flipper Length vs Bill Length by Species

Next, we make a scatter plot of flipper length vs bill length by penguin 
species below and saved the scatter plot to my project directory. 

```{r scatterplot}
plot_peng <- ggplot(penguins, aes(x = bill_length_mm,
                                  y = flipper_length_mm, color = species)) + 
             geom_point()

ggsave(scatter_penguins, plot = plot_peng, height = 4, width = 6)
```







