---
title: "p8105_hw1_mbc2178"
author: "Melvin Coleman"
date: "2022-09-23"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (tidyverse)
```

## Problem 1 

```{r}
data("penguins", package = "palmerpenguins")
```

The section below provides a brief description of the penguins dataset. 
The penguin dataset consists of the following summary statistics:
`r head(penguins, 5)`.
In this dataset, there are `r nrow(penguins) ` observations (rows) and 
`r ncol(penguins)` variables (columns). 
The mean flipper length in the penguins dataset is `r mean(pull(penguins, flipper_length_mm), na.rm=TRUE)` measured in millimeters(mm). 

## Scatter Plot of Flipper Length vs Bill Length by Species

Next, we made a scatter plot of `flipper_length_mm` vs `bill_length_mm` by penguin 
`species` below and saved the scatter plot to my project directory. 

```{r scatterplot}
plot_peng <- ggplot(penguins, aes(x = bill_length_mm,
                                  y = flipper_length_mm, color = species)) + 
             geom_point()

ggsave("penguins.pdf", height = 4, width = 6)
```

## Problem 2 

```{r, create_df}

peng_df <- tibble(
  norm_samp = rnorm (10, mean = 0),
  norm_samp_pos = norm_samp > 0,
  vec_hw = c("basketball", "handball", "streetball", "soccer","netball", 
              "dodgeball","football","kickball","softball","pongball"), 
  vec_factor_hw = factor(c("MARRIED","WIDOWED","SINGLE","MARRIED","SINGLE","WIDOWED","SINGLE","SINGLE","WIDOWED"
                            ,"MARRIED"),levels = c("MARRIED", "WIDOWED", "SINGLE"))
)
```
The mean of variable `norm_samp` is `r mean(pull(peng_df,norm_samp))`  
The mean of `norm_samp_pos` is `r mean(pull(peng_df, norm_samp_pos))` 
The mean of `vec_hw` is `r mean(pull(peng_df, vec_hw))` 
The mean of `vec_factor_hw` is `r mean(pull(peng_df, vec_factor_hw))` 

I noticed that trying to take the mean of the variables `vec_hw` and 
`vec_factor_hw` in the `peng_df` dataset yielded invalid results. In short, 
R couldn't perform the operation because the argument was not numeric or 
logical.

Nevertheless,I was able to calculate the mean of  variable `norm_samp` and 
`norm_samp_pos` in this dataset. 

Now let's apply the `as.numeric` function to our logical variable `norm_samp_pos`, 
character variable `vec_hw`, and factor variable `vec_factor_hw`. 

```{r convert_var, echo = FALSE}
  mean(as.numeric(pull(peng_df, norm_samp)))
  mean(as.numeric(pull(peng_df, norm_samp_pos)))
  mean(as.numeric(pull(peng_df, vec_hw)))
  mean(as.numeric(pull(peng_df, vec_factor_hw)))
```


















